sdmTMB: an R package for fast, flexible, and user-friendly generalized linear mixed effects models with spatial and spatiotemporal random fields

### 2023-03-11 doc/paper-jss/owls.Rmd
No need for extra optim now

```
#> ✓ Non-linear minimizer suggests successful convergence
#> ✓ Hessian matrix is positive definite
#> ✓ No extreme or very small eigenvalues detected
#> ✗ `ln_tau_E` gradient > 0.001
#>   See `?run_extra_optimization()`
#>   Or refit with `control = sdmTMBcontrol(newton_loops = 1)`
#> 
#> ✗ `ln_phi` gradient > 0.001
#>   See `?run_extra_optimization()`
#>   Or refit with `control = sdmTMBcontrol(newton_loops = 1)`
#> 
#> ✓ No fixed-effect standard errors are NA
#> ✓ No fixed-effect standard errors look unreasonably large
#> ✓ No sigma parameters are < 0.01
#> ✓ No sigma parameters are > 100
#> ✓ Range parameter doesn't look unreasonably large
```

We note that the magnitude of some log-likelihood gradients is slightly large for some fixed effects.

```{r}
fit_owl$gradients
```

We can run additional optimization steps with a Newton optimizer using the\
`run_extra_optimization()` function to further reduce these gradients.

```{r owl-newton, results='hide', cache=TRUE}
fit_owl2 <- run_extra_optimization(fit_owl, newton_loops = 1)
```

We can confirm that the gradients are smaller.

```{r sanity2, echo=TRUE}
sanity(fit_owl2)
```

@@

```{r print-fit-owls, warning=FALSE, echo=TRUE, message=FALSE}
print(fit_owl2)
```

@@

```{r tidy-owls}
tidy(fit_owl2, conf.int = TRUE)
tidy(fit_owl2, effects = "ran_pars", conf.int = TRUE)
```

@@

```{r predict-owls, message=FALSE, cache=TRUE}
pred <- predict(fit_owl2)

@@

```{r owls-resids}
set.seed(19208)
pred$resid <- residuals(fit_owl2)
```

@@

```{r zero-test-owls, echo=TRUE, message=FALSE}
s_nb2 <- simulate(fit_owl2, nsim = 400)
```

@@

```{r zero-test-owls3, echo=TRUE}
pred_fixed <- fit_owl2$family$linkinv(pred$est_non_rf)
r_nb2 <- DHARMa::createDHARMa(
  simulatedResponse = s_nb2,
  observedResponse = fit_owl2$data$count,
  fittedPredictedResponse = pred_fixed
)
DHARMa::testZeroInflation(r_nb2, plot = FALSE)
```

@@

We can then compress the annual predictions into mean values for plotting the spatially varying effect of NAO.
Most simply, we can extract the overall coefficient using the `b <- tidy(fit_owl2)` and add it to the spatially varying coefficients `zeta_s`.
However, this method does not give us spatially explicit estimates of uncertainty.
To do this, we can get the output from the joint precision matrix simulation draws by using the `predict()` function again, but this time indicating a number of simulation draws (e.g., `nsim = 200`) and what component we want estimated (`sims_var = "zeta_s"`). 
We also extract the same number of draws for estimates of the main effect of NAO using the `spread_sims()` function. 
The simulation draws for both can then be combined before calculating the median and lower and upper quantiles from these draws (in this case we will use 80%) and adding these values to our prediction data frame.

```{r p-mean, message=FALSE, echo=TRUE, eval=TRUE, cache=TRUE}
b <- tidy(fit_owl2, conf.int = TRUE)

nsim <- 200
zeta_s <- predict(fit_owl2, nsim = nsim, sims_var = "zeta_s")
sims <- spread_sims(fit_owl2, nsim = nsim)
beta <- sims$nao
combined <- beta + t(zeta_s)

>>>>

@@

```{r print-fit-owls, warning=FALSE, echo=TRUE, message=FALSE}
print(fit_owl)
```

@@

```{r tidy-owls}
tidy(fit_owl, conf.int = TRUE)
tidy(fit_owl, effects = "ran_pars", conf.int = TRUE)
```

@@

```{r predict-owls, message=FALSE, cache=TRUE}
pred <- predict(fit_owl)

@@

```{r owls-resids}
set.seed(19208)
pred$resid <- residuals(fit_owl)
```

@@

```{r zero-test-owls, echo=TRUE, message=FALSE}
s_nb2 <- simulate(fit_owl, nsim = 400)
```

@@

```{r zero-test-owls3, echo=TRUE}
pred_fixed <- fit_owl$family$linkinv(pred$est_non_rf)
r_nb2 <- DHARMa::createDHARMa(
  simulatedResponse = s_nb2,
  observedResponse = fit_owl$data$count,
  fittedPredictedResponse = pred_fixed
)
DHARMa::testZeroInflation(r_nb2, plot = FALSE)
```

@@

We can then compress the annual predictions into mean values for plotting the spatially varying effect of NAO.
Most simply, we can extract the overall coefficient using the `b <- tidy(fit_owl)` and add it to the spatially varying coefficients `zeta_s`.
However, this method does not give us spatially explicit estimates of uncertainty.
To do this, we can get the output from the joint precision matrix simulation draws by using the `predict()` function again, but this time indicating a number of simulation draws (e.g., `nsim = 200`) and what component we want estimated (`sims_var = "zeta_s"`). 
We also extract the same number of draws for estimates of the main effect of NAO using the `spread_sims()` function. 
The simulation draws for both can then be combined before calculating the median and lower and upper quantiles from these draws (in this case we will use 80%) and adding these values to our prediction data frame.

```{r p-mean, message=FALSE, echo=TRUE, eval=TRUE, cache=TRUE}
b <- tidy(fit_owl, conf.int = TRUE)

nsim <- 200
zeta_s <- predict(fit_owl, nsim = nsim, sims_var = "zeta_s")
sims <- spread_sims(fit_owl, nsim = nsim)
beta <- sims$nao
combined <- beta + t(zeta_s)
